---
title: "MHW plotting"
output:
  html_notebook: default
  pdf_document: 
    toc: false
    number_sections: no
  html_document: 
    toc: true
  word_document: 
    toc: true
editor_options: 
  chunk_output_type: inline
---

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(prompt=FALSE, comment=NA, message=FALSE, tidy=TRUE, error=FALSE, eval=TRUE)
```

#### **DATA EXTRACTION FROM PAWSEY/ LIBRARY PACKAGE USED**

Run this code chunk **only once** to read in the data file and load the library needed, you do not need to run this after you change variable.

```{r DataExtraction/LibraryPackage}
## Library needed
library(dplyr)
library(ggplot2)
library(heatwaveR)
library(patchwork)

## Data Extraction
library('aws.s3')
Sys.setenv('USE_HTTPS' = TRUE)

# if you do not have environment variable set in .env file, comment the readRenviron() function and uncomment the Sys.setenv() function below

readRenviron(".env")

# Sys.setenv(
#   'AWS_DEFAULT_REGION' = '', 
#   'AWS_S3_ENDPOINT' = 'projects.pawsey.org.au', 
#   'AWS_ACCESS_KEY_ID' = PUT YOUR ACCESS KEY ID, 
#   'AWS_SECRET_ACCESS_KEY' = PUT YOUR SECRET ACCESS KEY
# )

awss3Connect <- function(filename){
  bucket <- 'scevo-data'
  fetchedData <- aws.s3::s3read_using(FUN = utils::read.csv,
                                      check.names = FALSE,
                                      object = filename,
                                      bucket = bucket,
                                      filename = basename(filename),
                                      opts = list(
                                        base_url = Sys.getenv('AWS_S3_ENDPOINT'),
                                        region = Sys.getenv('AWS_DEFAULT_REGION'),
                                        key = Sys.getenv('AWS_ACCESS_KEY_ID'),
                                        secret = Sys.getenv('AWS_SECRET_ACCESS_KEY')))
  return(fetchedData)
}

Rawdata <- awss3Connect(filename = 'arms/wiski.csv')  
```

------------------------------------------------------------------------

------------------------------------------------------------------------

#### **FUNCTIONS USED**

Run this code chunk **only once** to run the functions to be used, you do not need to run this after you change variable.

```{r FunctionsUsed}
# Data Cleaning Functions
clean_data <- function(raw_data, swansites, depth) {
  samp_data <- raw_data %>%
    dplyr::filter(`Program Site Ref` %in% swansites, `Sample Depth (m)` <= depth) %>%
    dplyr::select(`Collect Date`, `Temperature (deg C)`)
  
  # Remove NA data
  samp_data <- na.omit(samp_data)
  
  # Change column names
  colnames(samp_data) <- c("t", "temp")
  
  # Change date format
  samp_data$t <- as.Date(samp_data$t, format = "%d/%m/%Y")
  
  return(samp_data)
}

average_data <- function(cleaned_data) {
  avg_temp <- cleaned_data %>%
  group_by(t) %>%
  summarise(avg_temperature = mean(temp))
}

climatologyperiod <- function() {
  note <- paste("Data available for site", swansites, "at depth <=", depth, "(m) are from", min(averaged_data$t), "to", max(averaged_data$t), ".")
  return(note)
}

# Data Interpolation Function
interpolate_data <- function(averaged_data) {
  # Create a sequence of daily dates
  daily_dates <- seq(min(averaged_data$t), max(averaged_data$t), by = "day")
  
  # Perform linear interpolation
  interpolated_data <- approx(averaged_data$t, averaged_data$avg_temperature, xout = daily_dates)
  
  # Convert to dataframe
  interpolated_df <- data.frame(date = interpolated_data$x, value = interpolated_data$y)
  
  # Change column names
  colnames(interpolated_df) <- c("t", "temp")
  
  return(interpolated_df)
}

# Event Detection Function
detect_events <- function(interpolated_data) {
  ts <- ts2clm(interpolated_data, climatologyPeriod = c(min(interpolated_data$t), max(interpolated_data$t)))
  mhw <- detect_event(ts)
  
  return(mhw)
}

# Data Visualization Functions
visualize_weekly_raw_data <- function(averaged_data, start_date_plot, end_date_plot) {
  filtered_data <- averaged_data[averaged_data$t >= start_date_plot & averaged_data$t <= end_date_plot,]
  footnote <- "* This plot shows the raw weekly sea surface temperature of the selected selected timeframe (START and END date)."
  ggplot(filtered_data, aes(x = t, y = avg_temperature)) +
    geom_point(color = "red", size = 2) + 
    labs(x = "", y = "Temperature [\u00B0C]", caption = footnote) +
    theme_bw() +
    theme(plot.caption = element_text(hjust = 0))
}

visualize_exceedance <- function(interpolated_df, threshold, start_date_plot, end_date_plot) {
  exc <- exceedance(interpolated_df, threshold = threshold)
  exc_thresh <- exc$threshold[exc$threshold$t >= start_date_plot & exc$threshold$t <= end_date_plot, ]
  footnote <- glue::glue(
    "* temp   :  daily sea surface temperature. \n",
    "* thresh :  the threshold of daily sea surface temperature for highlighting. \n", 
    "* The event highlighted is the event that exceeds the temperature threshold set within the\n",
    "   selected timeframe (START and END date of this plot), instead of the whole period of data.")
  
  ggplot(data = exc_thresh, aes(x = t)) +
    geom_flame(aes(y = temp, y2 = thresh, fill = "all"), show.legend = FALSE) +
    geom_line(aes(y = temp, colour = "temp")) +
    geom_line(aes(y = threshold, colour = "thresh"), size = 1.0) +
    scale_colour_manual(name = "Legend", values = c("temp" = "black", "thresh" = "forestgreen")) +
    scale_fill_manual(name = "Event Colour", values = c("all" = "salmon")) +
    guides(colour = guide_legend(override.aes = list(fill = NA))) +
    scale_x_date(date_labels = "%b %Y") +
    labs(x = "", y = "Temperature [\u00B0C]", caption = footnote) +
    theme_bw() +
    theme(plot.caption = element_text(hjust = 0))
}

visualize_event_line <- function(mhw, spread, metric, start_date_plot, end_date_plot, category) {
  footnote <- glue::glue(
    "* Temperature :  daily sea surface temperature. \n",
    "* Climatology   :  is calculated based on {min(averaged_data$t)} to {max(averaged_data$t)}. \n",
    "* Threshold      :  the 90th percentile (default) of daily sea surface temperature. \n", 
    "* The event highlighted is the largest event based on the selected timeframe (START and END date of this plot),\n",
    "   instead of the whole period of data.")
  messages <- "No events detected within the selected period. Please choose another date period."
  
  if(category) {
    plot <- tryCatch(
      event_line(mhw, spread = spread, metric = metric, start_date = start_date_plot, end_date = end_date_plot, category = TRUE),
      error = function(e) {
        message(messages)
        return(messages)
      }
    )
  } else {
    plot <- tryCatch(
      event_line(mhw, spread = spread, metric = metric, start_date = start_date_plot, end_date = end_date_plot),
      error = function(e) {
        message(messages)
        return(messages)
      }
    )
  }
  
  if(is.character(plot)) {
    return(plot)  # Return the message directly
  } else {
    plot <- plot + labs(caption = footnote) +
      theme(plot.caption = element_text(hjust = 0))
    return(plot)
  }
}

visualize_lolli_plot <- function(mhw, event_count, start_date_plot, end_date_plot) {
  filtered_events <- mhw$event[mhw$event$date_peak >= start_date_plot & mhw$event$date_peak <= end_date_plot, ]
  footnote <- glue::glue("* The event highlighted is the top ", event_count, " largest event based on the selected timeframe (START and END date of this plot),\n   instead of the whole period of data.")
  ggplot(filtered_events, aes(x = date_start, y = intensity_max)) +
  geom_lolli(colour = "salmon", colour_n = "red", n = 3) +
  labs(y = expression(paste("Max. intensity [", degree, "C]")), caption = footnote) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))
}

visualize_bubble_plot <- function(mhw, start_date_plot, end_date_plot) {
  filtered_events <- mhw$event[mhw$event$date_peak >= start_date_plot & mhw$event$date_peak <= end_date_plot, ]
  footnote <- ("* The event highlighted is the event that exceeds the temperature threshold set within the selected timeframe\n   (START and END date of this plot), instead of the whole period of data.")
  ggplot(filtered_events, aes(x = date_peak, y = intensity_max)) +
    geom_point(aes(size = intensity_cumulative), shape = 21, fill = "salmon", alpha = 0.8) +
    labs(x = NULL, y = "Maximum Intensity [°C]", size = "Cumulative Intensity [°C x days]", caption = footnote) +
    scale_size_continuous(range = c(0.1, 6), 
                          guide = guide_legend(title.position = "top", direction = "horizontal")) +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.key.size = unit(0.2, "lines"),
          legend.title = element_text(size=8), 
          legend.text = element_text(size=6),
          legend.box.background = element_rect(colour = "black"),
          plot.caption = element_text(hjust = 0))
}
```

------------------------------------------------------------------------

------------------------------------------------------------------------

#### **Variables to be changed (change the value and run the respective code chunk):**

<br/>

#### Step 1 (Select your site)

Select a site from: (**`ARM`**, **`BLA`**, **`FP1`**, **`FP7`**, **`FP22`**, **`CAV`**, **`HEA`**, **`NAR`**, **`NIL`**, **`STJ`**, **`MAY`**, **`RON`**, **`KIN`**, **`SUC`**, **`MEAD`**, **`MULB`**, **`WMP`**, **`REG`**, **`MSB`**, **`POL`**, **`SANDBR`**, **`KMO`**, **`JBC`**, **`VIT`**).

```{r SelectYourSite}
swansites <- c('ARM')
```

------------------------------------------------------------------------

#### Step 2 (Select your desired depth data)

Determine the **`depth`** measured you want to filter the data, the filtered data will only included data with depth less than or equal to the depth you set.

```{r FilterYourDepth}
depth = 2

# do not change the code below
cleaned_data = clean_data(Rawdata, swansites, depth) 
averaged_data = average_data(cleaned_data)
climatologyperiod()
```

------------------------------------------------------------------------

#### Step 3 (For all data visualization)

Determine **start and end date** for data visualization, this can only within the available date period outlined in **Step 2**.

```{r SelectTheTimeFrameForPlotting}
start_date_plot = "2010-04-12"
end_date_plot = "2012-12-12"
```

------------------------------------------------------------------------

#### Step 4 (For [event line](https://robwschlegel.github.io/heatwaveR/reference/event_line.html) plot)

Determine the **period (day) before & after the peak event** you want the graph to show (for event line plot); and determine if you want the plot to be **categorised**.

The **metric** tells the function how to choose the event that should be highlighted as the 'greatest' of the events in the chosen period. You may choose from the following options: **`intensity_mean`**, **`intensity_max`**, **`intensity_var`**,**`intensity_cumulative`**, **`intensity_mean_relThresh`**, **`intensity_max_relThresh`**, **`intensity_var_relThresh`**, **`intensity_cumulative_relThresh`**, **`intensity_mean_abs`**, **`intensity_max_abs`**, **`intensity_var_abs`**, **`intensity_cumulative_abs`**, **`rate_onset`**, **`rate_decline`**.

The plot may not necessarily correspond to the biggest event of the specified metric within the entire time series; instead, this plot shows the biggest event within the **`start_date_plot`** and **`end_date_plot`** you chose.

To plot the largest event within the whole time series, make sure you chose the **`start_date_plot`** and **`end_date_plot`** that cover the whole time series.

```{r CustomiseVariableForEventLinePlot}
spread = 180
category = FALSE
metric = "intensity_cumulative"
```

------------------------------------------------------------------------

#### Step 5 (For [exceedance](https://robwschlegel.github.io/heatwaveR/reference/exceedance.html) plot)

Determine temperature **`threshold`** to be detected (for exceedance plot).

```{r CustomiseThresholdToDetectExceedanceEvent}
threshold = 25
```

------------------------------------------------------------------------

#### Step 6 (For lolli plot)

Determine the **event count** to be highlighted (for lolli plot), the top x highest event will be in brighter red colour than others. The graph plotted here is adapted from the [**`lolli_plot()`**](https://robwschlegel.github.io/heatwaveR/reference/lolli_plot.html) function in the **heatwaveR package**, using **`ggplot()`**.

```{r CustomiseHighlightedEventForLolliPlot}
event_count = 3
```

------------------------------------------------------------------------

------------------------------------------------------------------------

#### **DATA TRANSFORMATION**

Run this code chunk to transform your data into the correct format to plot graphs, rerun this **every time you change your site and/or desired depth data**.

```{r DataTransformation}
cleaned_data = clean_data(Rawdata, swansites, depth) 
averaged_data = average_data(cleaned_data)
interpolated_data = interpolate_data(averaged_data) 
mhw = detect_events(interpolated_data)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

#### **PLOT GRAPHS**

Run these code chunks to plot the graphs respectively, rerun this **every time you change variable**:

##### [Weekly Raw Data]{.underline}

```{r WeeklyRawDataPlot}
visualize_weekly_raw_data(averaged_data, start_date_plot, end_date_plot) 
```

##### [Exceedance]{.underline}

```{r ExceedancePlot}
visualize_exceedance(interpolated_data, threshold, start_date_plot, end_date_plot)
```

##### [Event Line]{.underline}

```{r EventLinePlot}
visualize_event_line(mhw, spread, metric, start_date_plot, end_date_plot, category)
```

##### [Lolli Plot]{.underline}

```{r LolliPlot}
visualize_lolli_plot(mhw, event_count, start_date_plot, end_date_plot)
```

##### [Bubble Plot]{.underline}

```{r BubblePlot}
visualize_bubble_plot(mhw, start_date_plot, end_date_plot)
```
